---
title: "Nothing is zero in a multifactorial world (and why it matters)"
output:
  beamer_presentation:
    slide_level: 2
    keep_tex: true
    includes:
      in_header: ./header.tex
      after_body: ./suffix.tex
author: "Ben Bolker \\newline McMaster University"
date: "17 September 2021"
bibliography: discrete.bib
csl: reflist2.csl
---

<!-- tex hacks required: remove empty frame at beginning; break line in title (Burnham ref; add \framebreak manually to refs (ugh) -->
<!-- 
apa.csl is a slightly hacked version of APA 
  (modified for "et al" after 2 authors in text)
  -->
<!-- blockquote:
  https://css-tricks.com/snippets/css/simple-and-nice-blockquote-styling/ -->
<!-- center:
    https://www.w3schools.com/howto/howto_css_image_center.asp -->
<!-- .refs is style for reference page (small text) -->
<style>
.refs {
font-size: 14px;
}
.sm_block {
 font-size: 20px;
}
h2 { 
 color: #3399ff;		
}
h3 { 
 color: #3399ff;		
}
.title-slide {
   background-color: #55bbff;
   }
blockquote {
  background: #f9f9f9;
  border-left: 10px solid #ccc;
  margin: 1.5em 10px;
  padding: 0.5em 10px;
  quotes: "\201C""\201D"
}
blockquote:before {
  color: #ccc;
  content: open-quote;
  font-size: 4em;
  line-height: 0.1em;
  margin-right: 0.25em;
  vertical-align: -0.4em;
}
blockquote p {
  display: inline;
}
.center {
    display: block;
    margin-left: auto;
    margin-right: auto;
    width: 50%;
}
</style>
<!--    content: url(https://i.creativecommons.org/l/by-sa/4.0/88x31.png)
>
<!-- Limit image width and height -->
<style type="text/css">
img {     
  max-height: 560px;     
  max-width: 700px; 
}
div#before-column p.forceBreak {
	break-before: column;
}

div#after-column p.forceBreak {
	break-after: column;
}
</style>
```{r setup,echo=FALSE,message=FALSE, include=FALSE}
library(tidyverse)
library("ggplot2"); theme_set(theme_classic())
library("reshape2")
library("ggExtra")
library("MASS")
library("knitr")
opts_chunk$set(echo=FALSE,fig.width=4,fig.height=4, out.height="0.8 \\textheight")
```
## acknowledgements

money: NSERC

ideas: Jonathan Dushoff, Marm Kilpatrick, Brian McGill, Daniel Park, Daniel Turek

## what is a multifactorial system?

- many processes contribute to pattern
- quantify *how* each process affects the system,  
rather than testing *whether* we can detect its impact

This talk is mostly about *inferential* or *scientific* questions

## what are we trying to do?

- **prediction**: what do we expect in a specified scenario?
- **inference**: what factors are affecting observed outcomes, and how much?
   - (or: prediction/estimation of effects \cemph{with appropriate uncertainty bounds})
   
## why do we pretend some effects are zero?

we often consider \cemph{point hypotheses}, often as *counterfactual null hypotheses*
   - e.g.: "the delta variant has the same $R_0$ as earlier variants"; "masks do not prevent COVID transmission"
   - never\footnotemark true in biology/psychology/economics/epidemiology ...
   
more general we might want to test among \emph{discrete} hypotheses:

- "strong inference" [@platt_strong_1964]:  
(binary, discrete) alternative hypotheses 
- testing among *many* discrete hypotheses  
[@taper_evidential_2015]

\footnote[1]{what, never?}

## the method of multiple working hypotheses

- Chamberlin's \emph{method of multiple working hypotheses}  
[@raup_method_1995]

<blockquote>
... the measure of participation of each [process] must be determined before a satisfactory elucidation can be reached. The full solution therefore involves not only a recognition of multiple participation but an estimate of the measure and mode of each participation ...
</blockquote>

## discretomania

- null hypothesis testing as measure of \emph{clarity} (\bemph{OK} @dushoff_i_2019)
- "feature/variable selection" in data science (\bemph{OK but maybe not optimal?})
- stepwise regression (\cemph{not OK!})
- model averaging (OK?)

Why throw away information?

<!-- animation?

https://gist.github.com/mattblackwell/0d26d5c8f61f231570d61ccd62fe511f 
https://www.mattblackwell.org/blog/2021/04/06/rmd-overlays/

-->

## overfitting (1)

```{r overfit1}
## http://www.mathworks.com/products/matlab/demos.html?
##  file=/products/demos/shipping/matlab/census.html
dd <- data.frame(time = seq(1900,2000,by=10),
           pop = c(75.995,91.972,105.711,123.203,131.669,
                   150.697,179.323,203.212,226.505,249.633,281.422)) %>%
    mutate(sct=(time-1950)/50)
predframe <- tibble(time=seq(1900,2020, length=101)) %>% mutate(sct=(time-1950)/50)
orders = c(1:4,8)
pred <- purrr::map_dfr(setNames(orders, orders),
               ~ {
                   m <- lm(pop ~ poly(sct, ., raw=TRUE), data=dd)
                   data.frame(predframe,
                              pop = predict(m, newdata=predframe))
               },
               .id = "order")
ggplot(dd, aes(time, pop)) + geom_point(size=3) +
    geom_line(data=pred, aes(colour=order)) +
    coord_cartesian(ylim=c(0,350),
                    xlim=c(1890, 2025), expand=FALSE) +
    labs(x="year", y="US population (millions)")
```

## overfitting/underfitting {.columns-2 .build}

\begincols
\begincol{0.5\textwidth}

- \onslide<1-> avoid omitting important predictors
- \onslide<2-> avoid overfitting
- \onslide<3-> $\approx$ optimize \cemph{bias-variance tradeoff}
- \onslide<4-> any data-based approach to inclusion (stepwise etc.) runs afoul of the \emph{Texas sharpshooter fallacy}

\endcol
\begincol{0.5\textwidth}

\onslide<1->
<!-- \pause -->

\includegraphics[width=\textwidth]{pix/330px-The_Three_Bears_-_Project_Gutenberg_eText_17034.jpg}

\tiny Rackham 1837

\endcol
\endcols

## no free lunches

## multimodel averaging  (MMA) [@burnham_model_2002]

- ``information-theoretic" approach (Kullback-Leibler distance)
- 
- typical approach
    - fit full model; dredge; compute AIC weights
    - present table of models with $\Delta$AIC & weights
    - compute model-averaged point estimates and CIs
- **many** ways to do it wrong  
e.g. @cade_model_2015; @brewer_relative_2016; @galipaud_ecologists_2014
     - ... but let's not worry about that now

## why *not* use MMA?

three problems:

- conceptual
- computational
- inferential

## conceptual problem: discretization {.build .columns-2}

\begincols

\begincol{0.5\textwidth}

- information-theoretic (IT) approaches often framed as descendants of Chamberlin [@elliott_revisiting_2007]
- but IT focuses on differentiating **discrete** hypotheses/models
- submodels are always straw men
- ¿ does this actually cause problems ?

\endcol
\begincol{0.5\textwidth}

\pause
\includegraphics[width=\textwidth]{pix/eyam_ternary.png}

{\small Estimated contribution of plague transmission modes in Eyam 1665}

\endcol
\endcols

## computational problem: efficiency {.columns-2 .build}

\begincols
\begincol{0.5\textwidth}

\newcommand{\bigO}{{\mathcal{O}}}

- model average may mean fitting **lots** of models: $\bigO(2^K K^2 n)$
- we do have lots of computers ...

\endcol
\begincol{0.5\textwidth}

\pause

\includegraphics[width=0.75\textwidth]{pix/bitcoin_nyc.png}

\endcol
\endcols

## inferential problem: undercoverage {.build}

- various methods for constructing MMA CIs [@burnham_model_2002;@fletcher2012model;@kabaila_model-averaged_2016]
- MMA CIs are generally **too narrow**  
[@turek2013frequentist;@kabaila_model-averaged_2016;@dormann_model_2018] but cf. @burnham_model_2002
- general property of shrinkage estimators? e.g. ridge-regression CI width $\geq$ least-squares CI [@obenchain_classical_1977]
- **no free lunch**; can we ever gain certainty by shrinkage?

\pause

\includegraphics[width=0.7\textwidth]{pix/dormann_coverage.png}

@dormann_model_2018, Figure 5

## solutions: shrinkage estimators

- lasso/ridge/etc.; random forest etc.; Bayesian + regularizing priors
- $\bigO(K^2 n)$ or $\bigO(K n \log(n))$ (compare $\bigO(2^K K^2 n)$)
[@hardy_machine_2017; @louppe_understanding_2014]  
- may need to tune shrinkage by cross-validation
- but ...

## possible reactions

- "CIs only undercover a little bit"
    - then they're not good CIs; increased certainty is an illusion
- "I'm only interested in prediction"
    - really?
    - do you want reliable CIs on your predictions?
- "MMA is the only shrinkage estimator available for my problem"
    - let's work on this	
	
## conclusions: what should you do?

- for **inference**:
    - use the full model
    - *a priori* model reduction [@harrell_regression_2001]
- for **prediction**:
    - beware CIs from shrinkage estimates
    - help develop efficient, principled shrinkage estimation methods
    - use non-neutral, informative Bayesian priors?  
[@crome_novel_1996]

## References {.refs .columns-2 .allowframebreaks}

\tiny



